Adversarial Variational Optimization of\\ Non-Differentiable Simulators

### 2018-01-30 icml/avo.tex
add lfvi

This work sits at the intersection of several lines of research related to
likelihood-free inference, approximate Bayesian computation (ABC),
implicit generative models, and variational inference.
Viewed from the literature around implicit generative models based on neural networks,
the proposed method can be considered as a direct adaptation of
generative adversarial networks~\cite{goodfellow2014generative,2017arXiv170104862A, 2017arXiv170107875A} to
non-differentiable simulators using variational optimization~\cite{2012arXiv1212.4507S,staines2013optimization}.

@@

More generally, likelihood-free
inference is intimately tied to a class of algorithms that can be framed as
density estimation-by-comparison, as reviewed in \cite{2016arXiv161003483M}. In most cases, these
inference algorithms are formulated as an iterative two-step process where the
model distribution is first compared to the true data distribution and then
updated to make it more comparable to the latter. Relevant work in this
direction includes those that rely on a classifier to estimate the discrepancy between the observed data and the model distributions
\cite{gutmann2012noise,cranmer2015approximating,cranmer2016experiments,2016arXiv161110242D,gutmann2017likelihood,rosca2017variational}.
Also of direct relevance in the likelihood-free setup, Hamiltonian ABC~\cite{meeds2015hamiltonian}
estimate gradients with respect
to $\bftheta$ through finite differences from multiple forward passes of the
simulator with variance reduction strategies based on controlling the source of
randomness used for the latent variable $\bfz$.

@@

This work has also many connections to work on variational inference, in which the
goal is to optimize the recognition model $q(\bfz, \bftheta|\bfpsi)$ so that it
is close to the true posterior $p(\bfz, \bftheta |\mathbf{x})$. There have been efforts
to extend variational inference to intractable likelihoods; however, they
require restrictive assumptions.  In \cite{tran2017variational}, the authors
consider Variational Bayes with an Intractable Likelihood (VBIL). In that
approach ``the only requirement is that the intractable likelihood can be
estimated unbiasedly.'' In the case of simulators, they propose to use the
ABC-likelihood with an $\epsilon$-kernel. The ABC likelihood is only unbiased as
$\epsilon \to 0$,   thus this method inherits the drawbacks of the ABC-likelihood
including the choice of summary statistics and the inefficiency in evaluating
the ABC likelihood for high-dimensional data and small $\epsilon$.

>>>>

This work sits at the intersection of several lines of research related to
likelihood-free inference, approximate Bayesian computation (ABC),
implicit generative models, and variational inference.
Viewed from the literature around implicit generative models based on neural networks,
the proposed method can be considered as a direct adaptation of
generative adversarial networks~\cite{goodfellow2014generative,2017arXiv170104862A} to
non-differentiable simulators using variational optimization~\cite{2012arXiv1212.4507S}.

@@

More generally, likelihood-free
inference is intimately tied to a class of algorithms that can be framed as
density estimation-by-comparison, as reviewed in \cite{2016arXiv161003483M}. In most cases, these
inference algorithms are formulated as an iterative two-step process where the
model distribution is first compared to the true data distribution and then
updated to make it more comparable to the latter. Relevant work in this
direction includes those that rely on a classifier to estimate the discrepancy between the observed data and the model distributions
\cite{gutmann2012noise,cranmer2015approximating,cranmer2016experiments,2016arXiv161110242D,gutmann2017likelihood,rosca2017variational}.
% Also of direct relevance in the likelihood-free setup, Hamiltonian ABC~\cite{meeds2015hamiltonian}
% estimate gradients with respect
% to $\bftheta$ through finite differences from multiple forward passes of the
% simulator with variance reduction strategies based on controlling the source of
% randomness used for the latent variable $\bfz$.

@@

This work has also many connections to work on variational inference, in which
the goal is to optimize the recognition model $q(\bfz, \bftheta|\bfpsi)$ so that
it is close to the true posterior $p(\bfz, \bftheta |\mathbf{x})$. There have
been efforts to extend variational inference to intractable likelihoods;
however, many require restrictive assumptions.  In \cite{tran2017variational},
the authors consider Variational Bayes with an Intractable Likelihood (VBIL). In
that approach ``the only requirement is that the intractable likelihood can be
estimated unbiasedly.'' In the case of simulators, they propose to use the
ABC-likelihood with an $\epsilon$-kernel. The ABC likelihood is only unbiased as
$\epsilon \to 0$,   thus this method inherits the drawbacks of the
ABC-likelihood including the choice of summary statistics and the inefficiency
in evaluating the ABC likelihood for high-dimensional data and small $\epsilon$.
More recently, \cite{2017arXiv170208896T} adapted variational inference to
hierarchical implicit models defined on simulators. In this work, the authors
step around the intractable likelihoods by reformulating the optimization of the
ELBO in terms of a neural and differentiable approximation $r$ of the log-likelihood
ratio $\log \tfrac{p}{q}$.
